### 2024-03-23 content/chapter_1.tex
:pencil: Add wave scattering models
:pencil: Add SWIPT introduction and R-E tradeoff

	\begin{subsection}{Concept}
		\gls{ris} is commonly known as a planar surface involving numerous wave scattering elements (a.k.a. unit cells, reflective patches), whose amplitude and phase responses can be engineered in real-time to achieve a desired radiation pattern.
		It behaves like a delicate \gls{rf} mirror with adjustable curvature and orientation, which allows the incident signals to be focused and redirected in a particular direction.
		As shown in Fig. \ref{fg:ris_architecture}, its typical architecture consists of three stacked layers and a controller \cite{Wu2020}.
		The top layer is a two-dimensional array of scattering elements printed on a dielectric substrate.
		The elements directly interact with the impinging waves, which are usually fabricated from metamaterial or patch/dipole antennas with sub-wavelength dimension and spacing.
		The middle layer is a copper ground plate that provides voltage reference and avoids signal leakage.
		The bottom later is a circuit board that associate each element with adjustable components, such as varactor and \gls{pin} diodes \cite{Dai2020}.
		It also hosts a \gls{fpga} controller that controls the circuit and coordinate with transceivers in the network.
		By adjusting the scatter response of all elements, the \gls{ris} can effectively manipulate the wavefront for a constructive or destructive superposition and thus improve the ambient wireless environment.
		\begin{figure}
			\centering
			\includegraphics[width=0.9\textwidth]{ris_architecture.pdf}
			\caption{A typical architecture of \gls{ris}. Modified from \cite{Wu2020}.}
			\label{fg:ris_architecture}
		\end{figure}
	\end{subsection}

@@

	\begin{subsection}{Applications}
		The channel manipulation capability of \gls{ris} unlocks a wide range of unprecedented applications, such as signal enhancement \cite{Wu2019}, interference suppression \cite{Jiang2022}, blockage bypassing \cite{Ghatak2021}, coverage extension \cite{Zeng2021}, and security control \cite{Almohamad2020}.
		It also has the potential to convey additional information \cite{Ye2022}, compensate for the Doppler effect \cite{Basar2021}, transform frequency-selective channels into frequency-flat \cite{Arslan2022}, improve the spatial diversity for multi-antenna systems \cite{Ozdogan2020a}, and create artificial time diversity for multi-user orthogonal \cite{Yang2019} and non-orthogonal \cite{Chen2023} multiple accesses.

>>>>

	\begin{subsection}{Concept}
		\gls{ris} is commonly known as a planar surface involving numerous wave scattering elements (a.k.a. unit cells, reflective patches), whose amplitude and phase responses can be engineered in real-time to achieve a desired radiation pattern.
		It behaves like a delicate \gls{rf} mirror with adjustable curvature and orientation, which allows the incident signals to be focused and redirected in a particular direction.
		As shown in Fig. \ref{fg:ris_architecture}, its typical architecture consists of three stacked layers and a controller \cite{Wu2020}.
		The top layer is a two-dimensional array of scattering elements printed on a dielectric substrate.
		The elements directly interact with the impinging waves, which are usually fabricated from metamaterial or patch/dipole antennas with sub-wavelength dimension and spacing.
		The middle layer is a copper ground plate that provides voltage reference and avoids signal leakage.
		The bottom later is a circuit board that associate each element with adjustable components, such as varactor and \gls{pin} diodes \cite{Dai2020}.
		It also hosts a \gls{fpga} controller that controls the circuit and coordinate with transceivers in the network.
		By adjusting the scatter response of all elements, the \gls{ris} can effectively manipulate the wavefront for a constructive or destructive superposition and thus improve the ambient wireless environment.
		\begin{figure}
			\centering
			\includegraphics[width=0.9\textwidth]{ris_architecture.pdf}
			\caption{A typical architecture of \gls{ris}. The figure is modified from \cite{Wu2020}.}
			\label{fg:ris_architecture}
		\end{figure}
	\end{subsection}

@@

	\begin{subsection}{Applications}\label{sc:ris_applications}
		The channel manipulation capability of \gls{ris} unlocks a wide range of unprecedented applications, such as signal enhancement \cite{Wu2019}, interference suppression \cite{Jiang2022}, blockage bypassing \cite{Ghatak2021}, coverage extension \cite{Zeng2021}, and security control \cite{Almohamad2020}.
		It also has the potential to convey additional information \cite{Ye2022}, compensate for the Doppler effect \cite{Basar2021}, transform frequency-selective channels into frequency-flat \cite{Arslan2022}, improve the spatial diversity for multi-antenna systems \cite{Ozdogan2020a}, and create artificial time diversity for multi-user orthogonal \cite{Yang2019} and non-orthogonal \cite{Chen2023} multiple accesses.

### 2024-03-23 content/chapter_2.tex
:pencil: Add SWIPT introduction and R-E tradeoff

\chapter{Background and Literature Review}\label{ch:background}

\begin{section}{\glsfmtfull{ris}}
	\begin{subsection}{Programmable Metamaterials}
		Metamaterials refer to artificial structures engineered for unusual properties that may not be found in nature.
		The concept was initially proposed by Victor Veselago in 1967, who conjectured the existence of mediums with negative dielectric constant $\epsilon < 0$ and negative permeability $\mu < 0$ \cite{Veselago1968}.
		Such metamaterials are known as ``negative-index'' because the refraction index is defined as the \emph{negative} square root $n = - \sqrt{\epsilon \mu} < 0$, in order to be consistent with Maxwell's equations.
		It was not until 1999 that their feasibility was experimentally demonstrated by John Pendry at Imperial College using split-ring resonators \cite{Pendry1999}.
		Since then, metamaterials have attracted significant interests due to their counterintuitive properties, to name a few:
		\begin{itemize}
			\item \emph{Negative refraction:} As shown in Fig. \ref{fg:nim_refraction}, the incident and refracted rays stay at the same side of the normal axis \cite{Veselago1968}. This phenomenon is in contrast to the usual refraction but can still be predicted from Snell's law
			\begin{equation}
				\frac{\sin \theta_1}{\sin \theta_2} = n.
			\end{equation}
			It is worth mentioning that a generalized law of refraction and refraction has been proposed in \cite{Yu2011}, which has become a standard reference for the design and analysis of metamaterials.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-Index Material\label{fg:nim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_refraction.tex}
					}
				}
				\subfloat[Positive-Index Material\label{fg:pim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/pim_refraction.tex}
					}
				}
				\caption{Refraction in negative and positive-index materials. Incident and refracted rays stay at the same side of the normal axis in a negative-index material.}
			\end{figure}
			\item \emph{Opposite wave direction:} As shown in Fig. \ref{fg:nim_flows}, the wave vector and energy flow (indicated by the Poynting vector) are opposite to each other in a negative-index material \cite{Pendry2004}. This can be inferred from the electric field equation
			\begin{equation}
				\vec{E} = \vec{E}_0 \exp(\jmath k z - \jmath \omega t)
			\end{equation}
			where $k = k_0 n < 0$ is the wavenumber, $\vec{E}_0$ and $k_0$ are the free-space electric field and wavenumber reference, $z$ is the propagation distance, $\omega$ is the angular frequency, and $t$ is the time.
			Negative-index materials are thus also called ``left-hand'' because the propagation direction of the electric and magnetic fields can be determined by a left-hand rule.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-Index Material\label{fg:nim_flows}]{
					\resizebox{!}{2cm}{
							\input{assets/chapter_2/nim_flows.tex}
					}
				}
				\subfloat[Positive-Index Material\label{fg:pim_flows}]{
					\resizebox{!}{2cm}{
						\input{assets/chapter_2/pim_flows.tex}
					}
				}
				\caption{Wave and energy have opposite directions in a negative-index material.}
			\end{figure}
		\end{itemize}

@@

		\begin{figure}[H]
			\centering
			\subfloat[Negative-Index Material\label{fg:nim_focus}]{
				\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_focus.tex}
				}
			}
			\subfloat[Positive-Index Material\label{fg:pim_diverge}]{
				\resizebox{0.45\columnwidth}{!}{
					\input{assets/chapter_2/pim_diverge.tex}
				}
			}
			\caption{Refraction through metamaterials. For negative-index material, beams diverging from a point source is set in reverse and converges back to another point.}
		\end{figure}
		\begin{figure}[H]
			\centering
			\subfloat[Incident Wave\label{fg:wave_incident}]{
				\resizebox{!}{4.04cm}{
					\input{assets/chapter_2/wave_incident.tex}
				}
			}
			\subfloat[Reflected Wave\label{fg:wave_reflected}]{
				\resizebox{!}{4cm}{
					\input{assets/chapter_2/wave_reflected.tex}
				}
			}
			\caption{Reflection through metamaterials. Yellow dots represent scattering elements. Solid and dashed lines denote wavefronts and rays, respectively. The scattering elements work together to manipulate the phases of incident waves, resulting in a focused beam steered in the intended direction.}
		\end{figure}
		It is worth noticing that refraction and reflection are different implementations of \emph{passive beamforming} where the phase and amplitude of the ambient signal are altered by the metamaterial for a desired net effect.
		In the next subsection, we will some typical \gls{ris} scattering models and their physical architecture.
	\end{subsection}

@@

					z = \sum_{i=2,\text{even}}^\infty \beta_i \mathbb{E}\bigl\{ y^i(t) \bigr\},

@@

\begin{section}{\glsfmtfull{swipt}}\label{sc:swipt}
	\begin{subsection}{\glsfmtfull{r-e} Tradeoff}

	\begin{subsection}{Receiver Architectures}

>>>>

\chapter{Background}\label{ch:background}

\begin{section}{\glsfmtfull{ris}}
	\begin{subsection}{Programmable Metamaterials}
		Metamaterials refer to artificial structures engineered for unusual properties that may not be found in nature.
		The concept was initially proposed by Victor Veselago in 1967, who conjectured the existence of mediums with negative dielectric constant $\epsilon < 0$ and negative permeability $\mu < 0$ \cite{Veselago1968}.
		Such metamaterials are known as ``negative-index'' because the refraction index is defined as the \emph{negative} square root $n = - \sqrt{\epsilon \mu} < 0$, in order to be consistent with Maxwell's equations.
		It was not until 1999 that their feasibility was experimentally demonstrated by John Pendry at Imperial College using split-ring resonators \cite{Pendry1999}.
		Since then, metamaterials have attracted significant interests due to their counterintuitive properties, to name a few:
		\begin{itemize}
			\item \emph{Negative refraction:} As shown in Fig. \ref{fg:nim_refraction}, the incident and refracted rays stay at the same side of the normal axis \cite{Veselago1968}. This phenomenon is in contrast to the usual refraction but can still be predicted from Snell's law
			\begin{equation}
				\frac{\sin \theta_1}{\sin \theta_2} = n.
			\end{equation}
			It is worth mentioning that a generalized law of refraction and refraction has been proposed in \cite{Yu2011}, which has become a standard reference for the design and analysis of metamaterials.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-index material\label{fg:nim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_refraction.tex}
					}
				}
				\subfloat[Positive-index material\label{fg:pim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/pim_refraction.tex}
					}
				}
				\caption{Refraction in negative and positive-index materials. Incident and refracted rays stay at the same side of the normal axis in a negative-index material.}
			\end{figure}
			\item \emph{Opposite wave direction:} As shown in Fig. \ref{fg:nim_flows}, the wave vector and energy flow (indicated by the Poynting vector) are opposite to each other in a negative-index material \cite{Pendry2004}. This can be inferred from the electric field equation
			\begin{equation}
				\vec{E} = \vec{E}_0 \exp(\jmath k z - \jmath \omega t)
			\end{equation}
			where $k = k_0 n < 0$ is the wavenumber, $\vec{E}_0$ and $k_0$ are the free-space electric field and wavenumber reference, $z$ is the propagation distance, $\omega$ is the angular frequency, and $t$ is the time.
			Negative-index materials are thus also called ``left-hand'' because the propagation direction of the electric and magnetic fields can be determined by a left-hand rule.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-index material\label{fg:nim_flows}]{
					\resizebox{!}{2cm}{
							\input{assets/chapter_2/nim_flows.tex}
					}
				}
				\subfloat[Positive-index material\label{fg:pim_flows}]{
					\resizebox{!}{2cm}{
						\input{assets/chapter_2/pim_flows.tex}
					}
				}
				\caption{Wave and energy have opposite directions in a negative-index material.}
			\end{figure}
		\end{itemize}

@@

		\begin{figure}[H]
			\centering
			\subfloat[Negative-index material\label{fg:nim_focus}]{
				\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_focus.tex}
				}
			}
			\subfloat[Positive-index material\label{fg:pim_diverge}]{
				\resizebox{0.45\columnwidth}{!}{
					\input{assets/chapter_2/pim_diverge.tex}
				}
			}
			\caption{Refraction through metamaterials. For negative-index material, beams diverging from a point source is set in reverse and converges back to another point.}
		\end{figure}
		\begin{figure}[H]
			\centering
			\subfloat[Incident wave\label{fg:wave_incident}]{
				\resizebox{!}{4.04cm}{
					\input{assets/chapter_2/wave_incident.tex}
				}
			}
			\subfloat[Reflected wave\label{fg:wave_reflected}]{
				\resizebox{!}{4cm}{
					\input{assets/chapter_2/wave_reflected.tex}
				}
			}
			\caption{Reflection through metamaterials. Yellow dots represent scattering elements. Solid and dashed lines denote wavefronts and rays, respectively. The scattering elements work together to manipulate the phases of incident waves, resulting in a focused beam steered in the intended direction.}
		\end{figure}
		It is worth noticing that refraction and reflection are different implementations of \emph{passive beamforming} where the phase and amplitude of the ambient signal are altered by the metamaterial for a desired net effect.
		In the next subsection, we will some typical \gls{ris} scattering models and their physical architecture.
	\end{subsection}

@@

					z \triangleq \sum_{i=2,\text{even}}^\infty \beta_i \mathbb{E}\bigl\{ y^i(t) \bigr\},

@@

\begin{section}{\glsfmtfull{swipt}}\label{sc:swipt}
	\begin{subsection}{Introduction}
		\gls{wit} and \gls{wpt} have been treated separately over the past century and have made significant progress in their respective fields.
		Interestingly, electromagnetic waves carry information and energy simultaneously and the same signal can be used for communication and power transfer.
		The idea of \gls{swipt} was first proposed in 2008 \cite{Varshney2008} and has since attracted significant attention from both academia and industry.
		It is a promising solution to connect and energize trillions of low-power mobile devices, providing power at microwatt level and coverage up to tens of meters in a unified manner \cite{Clerckx2018b}.
		\gls{swipt} can also smoothly shift between the two extreme cases to fully exploit the \gls{rf} spectrum and network infrastructure.
		It is envisioned that future network providers will be able to offer a complete wireless solution including data and power services, which is essential for the upcoming intelligent era.

		One of the most important issues in \gls{swipt} is that the energy harvester requires a much higher received signal power (several orders of magnitude) than the information decoder \cite{Clerckx2019}.
		Since the channel \gls{rf}-to-\gls{rf} efficiency $\eta_2$ is the primary constraint on the overall power efficiency, how to combat the pathloss and fading effects has been recognized as a crucial research topic for \gls{wpt} \gls{swipt}.
		Fortunately, this issue can be effectively mitigated by introducing a \gls{ris} to the environment.
		By carefully tuning the scatter response of the \gls{ris} elements, one can potentially achieve the following benefits:
		\begin{itemize}
			\item \emph{Energy focusing:} The scattered waves can be steered towards the receivers or focused on a dedicated ``hotspot zone'' to increase the harvester input power level. This is also helpful to extend the coverage and improve the reliability of the energy link.
			\item \emph{Beam splitting:} Instead of transmitting one strong beam towards each user, the energy signal can be split into multiple weaker beams rerouted by the \gls{ris} to even out the spatial power distribution. This is useful to bypass physical obstacles (especially in high-frequency and large-scale networks) and reduce the health risk of radiation.
		\end{itemize}
		\gls{ris} can also be used to assist the information link by \gls{snr} enhancement and interference suppression, as mentioned in subsection \ref{sc:ris_applications}.
	\end{subsection}

	\begin{subsection}{\glsfmtfull{r-e} Tradeoff}
		Despite \gls{wit} and \gls{wpt} share many similarities, their difference in design objectives, system architectures, and practical constraints make a joint implementation of \gls{swipt} particularly challenging.
		Some preference of \gls{wit} and \gls{wpt} are inherently conflicting, for example:
		\begin{itemize}
			\item \emph{Waveform and modulation:} Under an average power constraint, \gls{wit} favors Gaussian signaling with maximum entropy distribution \cite{Cover2005} while \gls{wpt} prefers deterministic (unmodulated) multisine with higher \gls{papr} \cite{Trotter2009}.
			\item \emph{Channel:} In a \gls{mimo} scenario, \gls{wit} favors full-rank \gls{nlos} with high spatial diversity while \gls{wpt} prefers rank-deficient \gls{los} with high spatial correlation \cite{Wu2022a}.
			\item \emph{Receiver:} The power sensitivity is usually in the range of -40 to -80 dBm for information receivers and -10 to -30 dBm for energy harvesters \cite{Lu2015}.
		\end{itemize}
		Those disparities translate to a fundamental tradeoff between information and power transfer in \gls{swipt} systems, which is often quantified by a convex \emph{\gls{r-e} region}.
		\begin{equation}
			\mathcal{C}_\mathrm{R-E}(P) \triangleq \Bigl\{ (r,e) : 0 \le r \le \log(1 + \gamma), \ 0 \le e \le z \Bigr\},
		\end{equation}
		where $P$ is the average transmit power, $\gamma$ is the \gls{snr} at the information decoder, and $z$ defined in \eqref{eq:dc_power_nonlinear} is uniquely mapped to the harvested \gls{dc} power.
		Each point in this region corresponds to a \emph{rate-energy pair} achieved by a particular \emph{resource allocation scheme}.
		It is worth mentioning that different transceive strategies (e.g., waveform and receiver design) can lead to totally different \gls{r-e} regions (instead of different points in the same region), which motivates a joint optimization of the transmitter, channel, and receiver.
	\end{subsection}

	\begin{subsection}{System Architecture}

		\begin{subsubsection}{Information and Energy Flows}

		\end{subsubsection}

		\begin{subsubsection}{Receiver Architectures}

		\end{subsubsection}
		As shown in Figs. \subref*{fg:swipt_colocated} and \subref*{fg:swipt_separated}, \gls{swipt} can be classified into two categories based on the receiver configuration:
		% \begin{itemize}
		% 	\item \emph{\gls{swipt}:} Energy and information are simultaneously delivered in the downlink to co-located and/or separated information receiver(s) and energy receiver(s);
		% 	\item \emph{\gls{wpcn}:} Energy is delivered in the downlink to active device(s) that generate and transmit information-bearing signals in the uplink;
		% 	\item \emph{\gls{wpbc}:} Energy is delivered in the downlink to passive node(s) that reflect part of incoming waves and embed information in the uplink.
		% \end{itemize}
		\begin{figure}[H]
			\centering
			\subfloat[\gls{swipt} with co-located receiver]{
				\resizebox{7cm}{!}{
					\input{assets/chapter_2/swipt_colocated.tex}
				}
				\label{fg:swipt_colocated}
			}
			\subfloat[\gls{swipt} with separated receivers]{
				\resizebox{7cm}{!}{
					\input{assets/chapter_2/swipt_separated.tex}
				}
				\label{fg:swipt_separated}
			}
			\\
			\subfloat[Monostatic backscatter]{
				\resizebox{7cm}{!}{
					\input{assets/chapter_2/backcom_monostatic.tex}
				}
				\label{fg:backcom_monostatic}
			}
			\subfloat[Bistatic backscatter]{
				\resizebox{7cm}{!}{
					\input{assets/chapter_2/backcom_bistatic.tex}
				}
				\label{fg:backcom_bistatic}
			}
			\caption{
				Information and energy flows in different WIPT schemes. Dashed and dotted lines denote energy and information flows, respectively.
			}
			\label{fg:wipt_schemes}
		\end{figure}

### 2024-03-23 content/declaration.tex
:pencil: Add SWIPT introduction and R-E tradeoff

I hereby declare that the contents presented in this dissertation are original and have been carried out by myself under the guidance of my supervisor Prof. Bruno Clerckx.
Any work from other researchers, scholars, or sources have been properly cited and acknowledged.
The contents have not been submitted in whole or in part for consideration of any other degree or qualification in any academic institution.
I am aware of the ethical standards and academic integrity policies of Imperial College London, and I have adhered to these principles throughout the course of my study.
In signing this declaration, I affirm my commitment to academic honesty, intellectual integrity, and the pursuit of knowledge in the service of truth and understanding.

>>>>

I hereby declare that the contents presented in this dissertation are original and have been carried out by myself under the guidance of my supervisor Prof. Bruno Clerckx.
Any work from other researchers, scholars, or sources have been properly cited and acknowledged.
Figures, tables, and results presented in this dissertation are generated by myself, unless otherwise stated.
I would like to acknowledge the use of large language models for assistance in improving the clarity and readability of introduction and background sections of this thesis.
The contents have not been submitted in whole or in part for consideration of any other degree or qualification in any academic institution.
I am aware of the ethical standards and academic integrity policies of Imperial College London, and I have adhered to these principles throughout the course of study.
In signing this declaration, I affirm my commitment to academic honesty, intellectual integrity, and the pursuit of knowledge in the service of truth and understanding.
